### 2018-11-12 source/Ch2-techniques/calibration.tex
asdasda

\section{Calibration Methods in Radio Interferometry}\label{section.calibration}
\pg
In this section, we will discuss the implementation of interferometric array calibration. Our analysis is based on the RIME formalism, described in \cref{section.RIME}. One key metric of calibration quality is the \emph{dynamic range}, mentioned in Sec. \ref{section.clean}. High dynamic ranges mean that a high contrast has been obtained, and fainter sources can be reached. Here, because the difference between thermal noise and artefacts becomes relevant, we redefine dynamic range as follows
\begin{equation}\label{eq.DR}
DR = \frac{S_{max}}{S_{min},\max(\sigma_{thermal},\sigma_{artefacts})}
\end{equation}
where $S_{max}$ is the flux of the brightest source, $S_{min}$ the flux of the faintest source detectable above the noise, $\sigma_{thermal}$ the thermal noise in the image, and $\sigma_{artefacts}$ the noise associated with calibration artefacts. This gives a definition for dynamic range useful even in fields with a single source. We can thus use it as a metric for calibration quality specifically.

@@

\pg
First-generation calibration methods (1GC methods) consist of open-loop calibration. This relies entirely on instrument stability, and thus imposes significant design constraints on radio telescopes. It consists of briefly observing an external calibrator before and after each observation run to find calibration solutions for those calibrator observations, and then interpolate between them over the ``target" observation. %\footnote{For a concrete example, see \href{http://www.analog.com/en/analog-dialogue/articles/open-loop-calibration-techniques.html}{the open-loop calibration techniques page of analog.com}.}
%\pg
%Phase calibration in the 1GC era `proper' was not necessary, as engineers were capable of ensuring adequate phase stability in contemporary interferometers. Phase was thus calculated relative to a fixed frame of reference, usually the central antenna of a 3-antenna array. 
In RIME terms (see \cref{section.RIME}), this consists of solving for a very basic form of $\Gjones_p$:
\begin{equation}
\Gjones_p = a_p \bm{\mathrm{I}} % \begin{bmatrix} e^{2 \pi i \nu ( \phi_p - \phi_0)} \end{bmatrix}
\end{equation}
where $a_p$ is a complex constant solved for during open-loop calibration. By interpolating between the two calibrator observations, it is possible to have a linear time-dependent gain estimate, though it is not much more complex than the above. %, $\nu$ the observing frequency, $\phi$ the phase at an antenna, and $\phi_0$ the phase at the reference antenna.

\pg
While values for $a_p$ and $b_p$ can in theory be found for both autocorrelation and both crosscorrelations (e.g. XY and YX for linear receptors), low signal-to-noise means that in practice, a single set of values is solved for per antenna\footnote{This reduces calibration to solving only for the intensity gains: the data can then only be used for \emph{intensity mapping} (e.g. \citepads{1957IAUS....4..159J}). This practice therefore precludes polarimetry.}. With these techniques, one can achieve dynamic ranges of about 100:1 (\citepads{2010A&A...524A..61N}).

@@

\pg
Second-generation calibration methods (2GC methods) are defined by \emph{self-calibration}\footnote{On the discovery of self-calibration and its evolution in parallel to adaptive optics, see the chapter titled "The Almost Serendipitous Discovery of Self-Calibration'' in "\href{http://library.nrao.edu/public/collection/02000000000280.pdf}{\citep{serendipitous}}.}, commonly referred to as \emph{self-cal} (\citepads{1984ARA&A..22...97P}). As described in \cref{section.RIME.FullSky.CVZ}, this method can only be deployed if the brightness matrix of the sky is the same for all baselines\footnote{This is not to say that the brightness matrix must contain only point sources, but rather that moving our interferometer 500m to the East should not change its measured visibilities} (i.e. that we are not affected by direction-dependent effects).

@@

\pg
Self-calibration and adaptive optics are essentially the same thing, albeit meeting different constraints. Most notably, interferometric data is digital, allowing radio astronomers to perform their adaptive optics correction after the observation.

@@

\pg
The great advantage of radio interferometry over optics, however, is that we can \emph{iterate} over progressively improved source models. This is because interferometric data is digital, and because we record phase information.

@@

\pg
The crucial insight of a faceting approach is then to split the image into a set of \emph{facets}, solving for a set of calibration solutions for each of these facets. These solutions are then applied to the visibilities when mapping them onto the dirty image. The faceting in DDF is shown in \cref{fig.facets}. The facet distribution shown in \cref{fig.facets} encourages an even flux distribution in multiple facets, and ensures that no facet is so small that too little signal would be available within. As we can see from \cref{fig.DDcal.effect}, direction-dependent calibration can drastically improve the quality of wide-field images. It currently represents the state of the art in interferometric calibration.

>>>>

\section{Calibration Methods in Radio Interferometry}\label{section.calibration}
\pg
In this section, we will discuss the implementation of interferometric array calibration. Our analysis is based on the RIME formalism, described in \cref{section.RIME}. One key metric of calibration quality is the \emph{dynamic range}, mentioned in Sec. \ref{section.clean}. High dynamic ranges mean that a high contrast has been obtained, and fainter sources can be reached. Here, because the difference between thermal noise and artefacts becomes relevant, we redefine dynamic range as follows
\begin{equation}\label{eq.DR}
DR = \frac{S_{max}}{\max(S_{min},\sigma_{thermal},\sigma_{artefacts})}
\end{equation}
where $S_{max}$ is the flux of the brightest source, $S_{min}$ the flux of the faintest source detectable above the noise, $\sigma_{thermal}$ the thermal noise in the image, and $\sigma_{artefacts}$ the noise associated with calibration artefacts. This gives a definition for dynamic range useful even in fields with a single source. We can thus use it as a metric for calibration quality specifically.

@@

\pg
First-generation calibration methods (1GC methods) consist of open-loop calibration. This relies entirely on instrument stability, and thus imposes significant design constraints on radio telescopes. It consists of briefly observing an external calibrator before and after each observation run to find calibration solutions for those calibrator observations, and then interpolate between them over the ``target" observation. %\footnote{For a concrete example, see \href{http://www.analog.com/en/analog-dialogue/articles/open-loop-calibration-techniques.html}{the open-loop calibration techniques page of analog.com}.}
%\pg
%Phase calibration in the 1GC era `proper' was not necessary, as engineers were capable of ensuring adequate phase stability in contemporary interferometers. Phase was thus calculated relative to a fixed frame of reference, usually the central antenna of a 3-antenna array. 
In RIME terms (see \cref{section.RIME}), this consists of solving for a very basic form of $\Gjones_p$:
\begin{equation}
\Gjones_p = a_p \bm{\mathrm{I}} % \begin{bmatrix} e^{2 \pi i \nu ( \phi_p - \phi_0)} \end{bmatrix}
\end{equation}
where $a_p$ is a complex constant solved for during open-loop calibration and $\mathrm{I}$ is the unit matrix. By interpolating between the two calibrator observations, it is possible to have a linear time-dependent gain estimate.%, though it is not much more complex than the above. %, $\nu$ the observing frequency, $\phi$ the phase at an antenna, and $\phi_0$ the phase at the reference antenna.

\pg
While values for $a_p$ can in theory be found for both autocorrelation and both crosscorrelations (e.g. XY and YX for linear receptors), low signal-to-noise means that in practice, a single set of values is solved for per antenna\footnote{This reduces calibration to solving only for the intensity gains: the data can then only be used for \emph{intensity mapping} (e.g. \citepads{1957IAUS....4..159J}). This practice therefore precludes polarimetry.}. With these techniques, one can achieve dynamic ranges of about 100:1 (\citepads{2010A&A...524A..61N}).

@@

\pg
Second-generation calibration methods (2GC methods) are defined by \emph{self-calibration}\footnote{On the discovery of self-calibration and its evolution in parallel to adaptive optics, see the chapter titled "The Almost Serendipitous Discovery of Self-Calibration'' in "\href{http://library.nrao.edu/public/collection/02000000000280.pdf}{\citep{serendipitous}}.}, commonly referred to as \emph{self-cal} (\citetads{1984ARA&A..22...97P}). As described in \cref{section.RIME.FullSky.CVZ}, this method can only be deployed if the brightness matrix of the sky is the same for all baselines\footnote{This is not to say that the brightness matrix must contain only point sources, but rather that moving our interferometer 500m to the East should not change its measured visibilities} (i.e. that we are not affected by direction-dependent effects).

@@

\pg
Self-calibration and adaptive optics are essentially the same thing: an attempt to correct for phase errors in the signal measured by an instrument, which have been introduced by the atmosphere (i.e. correct time-varying and direction-dependent processes). This is shown in \cref{fig.selfcal}. Self-calibration and adaptive optics nevertheless meet different constraints: most notably, because interferometric data is digital, radio astronomers can perform their corrections after the observation.

@@

\pg
The great advantage of radio interferometry over optics, however, is that we can \emph{iterate} over progressively improved source models. This is because we have the ability to record phase information.

@@

\pg
The crucial insight of a faceting approach is then to split the image into a set of \emph{facets}, solving for a set of calibration solutions for each of these facets. These solutions are then applied to the visibilities when mapping them onto the dirty image. The faceting performed by DDF is shown in \cref{fig.facets}. The facet distribution shown in \cref{fig.facets} encourages an even flux distribution in multiple facets, and ensures that no facet is so small that too little signal would be available within. As we can see from \cref{fig.DDcal.effect}, which is shows two cutouts from \cref{fig.facets} before and after performing direction-dependent calibration, this process can drastically improve the quality of wide-field images. It currently represents the state of the art in interferometric calibration.
